---
title: "p8105_hw6_mbc2178"
author: "Melvin Coleman"
date: "11/23/2022"
output: github_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)

set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1 



### Problem 2 

Let's examine the Washington Post's data on homicides in 50 large U.S. cities. 
We will refer to this raw dataset as `homicide_df`.

```{r}
homicide_df =
  read_csv('data/homicide-data.csv', col_names = TRUE) 
```

The `homicide_df` contains data from 50 large cities in the U.S. with `r ncol(homicide_df)`
fields or variables and `r nrow(homicide_df)` records or observations. This dataset includes
first and last names as well as ethnicity of homicide victims (`victim_last`,`victim_first`,
`victim_race`), dates of homicide report (`reported_date`) and the status of cases characterized
as closed with arrest or closed without arrest (`disposition`). In addition, the 
dataset also contains the longitude (`lon`) and latitude (`lat`) of the cities 
where the homicide occurred. 

Let's perform some data cleaning and manipulation on `homicide_df` using the 
`janitor::clean_names()` and `mutate`. We ensured that `victim_age` is a numeric 
variable by using an `if_else` statement that would apply NA to all missing values.

New variables `city_state` and a binary variable `homicide_solved` with 0 and 1 to 
indicate whether the homicide is solved were created. In addition, we limit our 
analysis for those whose `victim_race` is white or black and exclude the following 
cities Dallas,TX,Phoenix, AZ, and Kansas City, MO because these don’t report victim 
race. Also, we omit Tulsa,AL because it is a data entry mistake.

```{r}
homicide_df =
  homicide_df %>% 
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city,',',state),
    homicide_solved = if_else(disposition %in% c("Closed by arrest"), 1, 0),
    victim_age = ifelse(victim_age > 0, as.numeric(victim_age), NA)) %>% 
    
    filter(!city_state %in% c("Dallas,TX", "Phoenix, AZ", "Kansas City, MO","Tulsa,AL"),
           victim_race %in% c("White", "Black")) 
  
    
```

Let's now fit a logistic regression with resolved vs unresolved as the outcome and
victim age, sex and race as predictors for the city of Baltimore, MD. 
